Triangulating iconicity: Coding analysis

### 2024-06-15 paper_02_main_analyses.Rmd
knit 'n commit

@@

#panel B: ratings by semantic domain
pB <- d |>
  filter(category != "Other") |>
  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D",alpha=0.8) +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))

@@

@@

### Figure 5: Ratings and cumulative iconicity

@@

# panel A: ratings by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5) +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

@@

>>>>

# variant with shadow for use in paper
# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  ggfx::with_shadow(
    geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center"),
    sigma=0,
    x_offset=1.4,
    y_offset=1.4,
    colour="black") +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  ggfx::with_shadow(
    geom_half_point(),
    sigma=0,
    x_offset=1.4,
    y_offset=1.4,
    colour="black") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

#plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
#ggsave("figures/fig3-panelAB-shadow.png",height=4,width=9,bg="white")

@@

#panel B: ratings by semantic domain
pB <- d |>
  filter(category != "Other") |>
  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  geom_half_point(show.legend=F) +
  scale_colour_viridis_d(option="D") +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))

@@

# variant with shadow for possible use in paper

#panel B: ratings by semantic domain
pB <- d |>
  filter(category != "Other") |>
  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
  theme_tufte(base_size=16) +
  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
  ggfx::with_shadow(
    geom_half_point(show.legend=F),
    sigma=0,
    x_offset=1.4,
    y_offset=1.4,
    colour="black"
    ) +
  scale_colour_viridis_d(option="D") +
  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
  "Texture","Colour/Visual"))

#plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.6,2))
#ggsave("figures/fig4-panel_ratings-shadow.png",height=4,width=9,bg="white")

@@

### Figure 6: Ratings and cumulative iconicity

@@

# panel A: ratings by study and cumulative iconicity
pA <- ggplot(data=dp, 
             aes(x="",
                 y=rating,
                 colour=C_cumulative,
                 fill=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5) +
  NULL

pB <- ggplot(data=dp, 
             aes(x=C_cumulative,
                 y=rating,
                 colour=C_cumulative,
                 fill=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

@@

R1 asks for a variant of this figure where the dots have black outline, to make the most yellow dots more easily visible. A second variant with subtle dropshadow may be even better.

```{r R1_request}

pA <- ggplot(data=dp, 
             aes(x="",
                 y=rating,
                 colour=C_cumulative,
                 fill=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5,
                shape=21,
                colour="black") +
  NULL

pB <- ggplot(data=dp, 
             aes(x=C_cumulative,
                 y=rating,
#                 colour=C_cumulative,
                 fill=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point(shape=21,colour="black") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures/fig6-panelAB_ratings_variant1.png",height=5,width=9,bg="white")

# what about a drop shadow?

pA <- ggplot(data=dp, 
             aes(x="",
                 y=rating,
                 colour=C_cumulative,
                 fill=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  ggfx::with_shadow(
    geom_beeswarm(cex=2.5),
    sigma=0,
    x_offset=1.4,
    y_offset=1.4,
    colour="black"
  ) +
  NULL

pB <- ggplot(data=dp, 
             aes(x=C_cumulative,
                 y=rating)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot(aes(colour=C_cumulative)) +
  ggfx::with_shadow(
    geom_half_point(aes(colour=C_cumulative,fill=C_cumulative),
                    shape=21),
    sigma=0,
    x_offset=1.4,
    y_offset=1.4,
    colour="black") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures/fig6-panelAB_ratings_variant2.png",height=5,width=9,bg="white")
